\textbf{ Normalized Gaussian Processes }

### 2020-11-27 tex/ms.tex
ci

The denominator in the expressions above makes direct evaluation of the expectations

@@

    \frac{(n + 1) g_{n} \bar{\sigma}^\frac{n+2}{2}}{\mu^{n+2}} \, \mathbf{j} \, \mathbf{j}^\top

@@

    \frac{(n + 1) K \, g_{n} \bar{\sigma}^\frac{n}{2}}{\mu^{n+1}} \, \bar{\pmb{\sigma}} \, \mathbf{j}^\top

@@

    n \,  \bar{\sigma}^\frac{n - 2}{2} \bar{\pmb{\sigma}} \, \bar{\pmb{\sigma}}^\top + \bar{\sigma}^\frac{n}{2} \pmb{\Sigma}

@@

    \bar{\pmb{\sigma}}

@@

    \bar{\sigma}

@@

is the average of all the elements in $\pmb{\Sigma}$.

@@

    \frac{\alpha \, \left(\mathbf{s} \, \mathbf{s}^\top - \bar{\pmb{\sigma}} \, \bar{\pmb{\sigma}}^\top\right)}{\mu^2 \bar{\sigma}}  +
    \frac{\beta \, \left(\mathbf{s} \, \mathbf{s}^\top\right)}{\mu^2 \bar{\sigma}}

@@

    \bar{\sigma} \, \mathbf{j} - \bar{\pmb{\sigma}}

@@

    z & \equiv \frac{\bar{\sigma}}{\mu^2}

@@

@@

In the following sections, we empirically show that this interpretation is
correct, and the expression for the normalized covariance
(Equation~\ref{eq:SigmaTilde}) is accurate provided $z \ll 1$.

@@

                     & = 3 \, \bar{\sigma} \, \mathbf{j} \, \mathbf{j}^\top

@@

    where $\bar{\sigma}$ is the average of all entries in $\pmb{\Sigma}$ and

@@

        3 \, \bar{\sigma}^2 \mathbf{j} \, \mathbf{j}^\top

@@

            g_n (n + 1) \, \bar{\sigma}^\frac{n}{2} & n \, \mathrm{even}

@@

    \subsection{The $\mathbf{R}$ expectation}
    %
    The first several terms in $\mathbf{R}$ are given by
    %
    \setlength{\abovedisplayskip}{1em}
    \begin{align}
        \mathbf{R}_0 & = \mathbb{E}\left[ \mathbf{L} \, \mathbf{u} \, \mathbf{u}^\top \mathbf{L}^\top \right]
        \nonumber                                                                                                                                                              \\
                     & = \mathbf{L} \, \mathbb{E}\left[  \mathbf{u} \, \mathbf{u}^\top \right] \, \mathbf{L}^\top
        \nonumber                                                                                                                                                              \\
                     & = \pmb{\Sigma}
        %
        \\[1em]
        %
        \mathbf{R}_1 & = \frac{1}{K}\mathbb{E}\left[ \left(\mathbf{j}^\top \mathbf{L} \, \mathbf{u}\right) \mathbf{L} \, \mathbf{u} \, \mathbf{u}^\top \mathbf{L}^\top \right]
        \nonumber                                                                                                                                                              \\
                     & = \mathbf{0}
        %
    \end{align}
    %
    since its elements are all products of three standard normal random variables. Next,
    %
    \begin{align}
        %
        \mathbf{R}_2 & = \mathbb{E}\left[ \left(\mathbf{j}^\top \mathbf{L} \, \mathbf{u}\right)^2 \mathbf{L} \, \mathbf{u} \, \mathbf{u}^\top \mathbf{L}^\top \right]
        \nonumber                                                                                                                                                     \\
                     & = \xxx{...}
        \nonumber                                                                                                                                                     \\
                     & = 2 \pmb{\Sigma} \mathbf{j} \, \mathbf{j}^\top \pmb{\Sigma} + \mathrm{sum}(\pmb{\Sigma}) \pmb{\Sigma}
    \end{align}
    %
    \xxx{By trial and error, I found that}
    %
    \begin{align}
        %
        \mathbf{R}_4 & = \frac{5}{K^4}\mathbb{E}\left[ \left(\mathbf{j}^\top \mathbf{L} \, \mathbf{u}\right)^4 \mathbf{L} \, \mathbf{u} \, \mathbf{u}^\top \mathbf{L}^\top \right]
        \nonumber                                                                                                                                                                  \\
                     & = \xxx{...}
        \nonumber                                                                                                                                                                  \\
                     & = \frac{5}{K^4} \Big( 12 (\pmb{\Sigma} \mathbf{j} \, \mathbf{j}^\top)^2 \pmb{\Sigma} + 3 \, \mathrm{sum}(\pmb{\Sigma})^2 \pmb{\Sigma} \Big)
    \end{align}
    %
    Based on this pattern, the general term in $\mathbf{R}$ is
    %
    \begin{align}
        \mathbf{R}_n & =
        \begin{cases}
            \mathbf{0}                                                                                                                                                      & n \, \mathrm{odd}
            \\
            \frac{n (n + 1) g_{n}}{K^n} \, (\pmb{\Sigma} \, \mathbf{j} \, \mathbf{j}^\top)^\frac{n}{2}\pmb{\Sigma} + (n + 1) g_{n} \, \bar{\sigma}^\frac{n}{2} \pmb{\Sigma} & n \, \mathrm{even}
            \quad.
        \end{cases}
    \end{align}

    \subsection{The final result}
    Inserting the matrices $\mathbf{P}$, $\mathbf{Q}$, and $\mathbf{R}$ into the expression
    for $\tilde{\pmb{\Sigma}}$ and rearranging,
    we obtain
    %
    \begin{align}
        \tilde{\pmb{\Sigma}}
         & =
        \frac{1}{\mu^2} \pmb{\Sigma}
        +
        \frac{1}{\mu^2}
        \sum\limits_{n=1}^\infty
        \frac{(-1)^n(n + 1)}{\mu^{n}}
        \,
        \bigg[
            g_n \, \bar{\sigma}^\frac{n}{2} \, (\pmb{\Sigma} + \mu^2\mathbf{j} \, \mathbf{j}^\top)
            \, +
            \nonumber                 \\[0.5em]
         & \phantom{XXXXXXXXXXXXXXX.}
        K g_{n+1} \mu  \,
        \left(
        \left(\pmb{\Sigma} \, \frac{\mathbf{j} \, \mathbf{j}^\top}{K^2}\right)^\frac{n + 1}{2}
        +
        \left(\frac{\mathbf{j} \, \mathbf{j}^\top}{K^2} \, \pmb{\Sigma}\right)^\frac{n + 1}{2}
        \right)
        \, +
        \nonumber                     \\[0.5em]
         & \phantom{XXXXXXXXXXXXXXX.}
        n g_n \, \left(\pmb{\Sigma} \, \frac{\mathbf{j} \, \mathbf{j}^\top}{K^2}\right)^\frac{n}{2}\pmb{\Sigma}
        \bigg]
        \quad.
    \end{align}

>>>>

The denominators in the expressions above make direct evaluation of the expectations

@@

    \frac{(n + 1) g_{n} m^\frac{n+2}{2}}{\mu^{n+2}} \, \mathbf{j} \, \mathbf{j}^\top

@@

    \frac{(n + 1) K \, g_{n} m^\frac{n}{2}}{\mu^{n+1}} \, \mathbf{m} \, \mathbf{j}^\top

@@

    n \,  m^\frac{n - 2}{2} \mathbf{m} \, \mathbf{m}^\top + m^\frac{n}{2} \pmb{\Sigma}

@@

    \mathbf{m}

@@

    m

@@

is the average of all elements in $\pmb{\Sigma}$.

@@

    \frac{\alpha \, \left(\mathbf{s} \, \mathbf{s}^\top - \mathbf{m} \, \mathbf{m}^\top\right)}{\mu^2 m}  +
    \frac{\beta \, \left(\mathbf{s} \, \mathbf{s}^\top\right)}{\mu^2 m}

@@

    m \, \mathbf{j} - \mathbf{m}

@@

    z & \equiv \frac{m}{\mu^2}

@@

equal to the average element of the covariance matrix divided by the square of the
GP mean.

@@

In the following sections, we empirically show that this interpretation is
correct, and that the expression for the normalized covariance
(Equation~\ref{eq:SigmaTilde}) is accurate provided $z \ll 1$.

@@

                     & = 3 \, m \, \mathbf{j} \, \mathbf{j}^\top

@@

    where $m$ is the average of all entries in $\pmb{\Sigma}$ and

@@

        3 \, m^2 \mathbf{j} \, \mathbf{j}^\top

@@

            g_n (n + 1) \, m^\frac{n}{2} & n \, \mathrm{even}

@@

    \subsection{The $\mathbf{R}$ expectation}
    %
    The first several terms in $\mathbf{R}$ are given by
    %
    \setlength{\abovedisplayskip}{1em}
    \begin{align}
        \mathbf{R}_0 & = \mathbb{E}\left[ \mathbf{L} \, \mathbf{u} \, \mathbf{u}^\top \mathbf{L}^\top \right]
        \nonumber                                                                                                                                                              \\
                     & = \mathbf{L} \, \mathbb{E}\left[  \mathbf{u} \, \mathbf{u}^\top \right] \, \mathbf{L}^\top
        \nonumber                                                                                                                                                              \\
                     & = \pmb{\Sigma}
        %
        \\[1em]
        %
        \mathbf{R}_1 & = \frac{1}{K}\mathbb{E}\left[ \left(\mathbf{j}^\top \mathbf{L} \, \mathbf{u}\right) \mathbf{L} \, \mathbf{u} \, \mathbf{u}^\top \mathbf{L}^\top \right]
        \nonumber                                                                                                                                                              \\
                     & = \mathbf{0}
        %
    \end{align}
    %
    since its elements are all products of three standard normal random variables. Next,
    %
    \begin{align}
        %
        \mathbf{R}_2 & = \mathbb{E}\left[ \left(\mathbf{j}^\top \mathbf{L} \, \mathbf{u}\right)^2 \mathbf{L} \, \mathbf{u} \, \mathbf{u}^\top \mathbf{L}^\top \right]
        \nonumber                                                                                                                                                     \\
                     & = \xxx{...}
        \nonumber                                                                                                                                                     \\
                     & = 2 \pmb{\Sigma} \mathbf{j} \, \mathbf{j}^\top \pmb{\Sigma} + \mathrm{sum}(\pmb{\Sigma}) \pmb{\Sigma}
    \end{align}
    %
    \xxx{By trial and error, I found that}
    %
    \begin{align}
        %
        \mathbf{R}_4 & = \frac{5}{K^4}\mathbb{E}\left[ \left(\mathbf{j}^\top \mathbf{L} \, \mathbf{u}\right)^4 \mathbf{L} \, \mathbf{u} \, \mathbf{u}^\top \mathbf{L}^\top \right]
        \nonumber                                                                                                                                                                  \\
                     & = \xxx{...}
        \nonumber                                                                                                                                                                  \\
                     & = \frac{5}{K^4} \Big( 12 (\pmb{\Sigma} \mathbf{j} \, \mathbf{j}^\top)^2 \pmb{\Sigma} + 3 \, \mathrm{sum}(\pmb{\Sigma})^2 \pmb{\Sigma} \Big)
    \end{align}
    %
    Based on this pattern, the general term in $\mathbf{R}$ is
    %
    \begin{align}
        \mathbf{R}_n & =
        \begin{cases}
            \mathbf{0}                                                                                                                                           & n \, \mathrm{odd}
            \\
            \frac{n (n + 1) g_{n}}{K^n} \, (\pmb{\Sigma} \, \mathbf{j} \, \mathbf{j}^\top)^\frac{n}{2}\pmb{\Sigma} + (n + 1) g_{n} \, m^\frac{n}{2} \pmb{\Sigma} & n \, \mathrm{even}
            \quad.
        \end{cases}
    \end{align}

    \subsection{The final result}
    Inserting the matrices $\mathbf{P}$, $\mathbf{Q}$, and $\mathbf{R}$ into the expression
    for $\tilde{\pmb{\Sigma}}$ and rearranging,
    we obtain
    %
    \begin{align}
        \tilde{\pmb{\Sigma}}
         & =
        \frac{1}{\mu^2} \pmb{\Sigma}
        +
        \frac{1}{\mu^2}
        \sum\limits_{n=1}^\infty
        \frac{(-1)^n(n + 1)}{\mu^{n}}
        \,
        \bigg[
            g_n \, m^\frac{n}{2} \, (\pmb{\Sigma} + \mu^2\mathbf{j} \, \mathbf{j}^\top)
            \, +
            \nonumber                 \\[0.5em]
         & \phantom{XXXXXXXXXXXXXXX.}
        K g_{n+1} \mu  \,
        \left(
        \left(\pmb{\Sigma} \, \frac{\mathbf{j} \, \mathbf{j}^\top}{K^2}\right)^\frac{n + 1}{2}
        +
        \left(\frac{\mathbf{j} \, \mathbf{j}^\top}{K^2} \, \pmb{\Sigma}\right)^\frac{n + 1}{2}
        \right)
        \, +
        \nonumber                     \\[0.5em]
         & \phantom{XXXXXXXXXXXXXXX.}
        n g_n \, \left(\pmb{\Sigma} \, \frac{\mathbf{j} \, \mathbf{j}^\top}{K^2}\right)^\frac{n}{2}\pmb{\Sigma}
        \bigg]
        \quad.
    \end{align}
